---
title: "Figures"
output:
  word_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---

Script for plotting the figures in the paper

# Loading the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height=8, fig.width=16,message=FALSE,echo=FALSE, warnings = FALSE)
options(scipen = 3)
```

```{r}
library(magrittr)
library(broom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(purrr)
library(GGally)
library(moments)
library(lme4)
library(scales)
library(gtable)
library(grid)

setwd('C:/Users/Admin/Documents/Paper/MatingDesignVA-main')

p_df <- readRDS(file = 'output/Appendix_A.RData')

summary_df <- readRDS('output/summary_df.RData') 

ANOVA_df <- readRDS('output/ANOVA_df.RData')

expectation_df <- readRDS(file = 'output/figS1_expectation.RData') 

simulation_df <- readRDS(file = 'output/figS1_simulation.RData')

D_ancestral <- readRDS(file = 'output/D_ancestral.RData')

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


# Figures

## Fig. 1

Simulation scheme

## Fig. 2

```{r}
cross_df <- tibble(P_num = c(2, 4, 8, 16)) %>% 
  mutate(P = factor(P_num, levels = c(2, 4, 8, 16),
                    labels = paste0('P=', c(2, 4, 8, 16))),
         DC = P_num/2,
         FC = (P_num/2)**2,
         HC = (P_num * (P_num - 1)) / 2) %>% 
  pivot_longer(DC:HC, names_to = 'MD',
               values_to = 'crosses') %>% 
  mutate(MD = factor(MD,
                     levels = c('DC', 'FC', 'HC'))) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2'))


gtable_filter_remove <- function (x, name, trim = TRUE){
    matches <- !(x$layout$name %in% name)
    x$layout <- x$layout[matches, , drop = FALSE]
    x$grobs <- x$grobs[matches]
    if (trim) 
        x <- gtable_trim(x)
    x
}

elite <- summary_df %>% 
  filter(N == 'N=1000',
         G == 'G1_DH',
         P %in% paste0('P=', c(2, 4, 8, 16)),
         parameter == 'E_Vg',
         L == 'L=1000',
         ancestral == 'Elite',
         !(MD %in% c('DC', 'HC') & P == 'P=2')) %>% 
  ggplot(aes(x=MD, y = Mean))+
  geom_col(position = 'dodge', fill = cbPalette[1])+
  geom_text(y = 500, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_y_continuous(limits = c(0, 500))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  ylab(bquote('Estimated expectation for'~V[A]))+
  theme_light(base_size = 10)+
  theme(strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'none',
        axis.title.x = element_blank())

plot_tab_elite <- ggplotGrob(elite)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_elite, 
                     "axis-b-1",
                     trim=FALSE)

elite <- ggpubr::as_ggplot(tmp)

landrace <- summary_df %>% 
  filter(N == 'N=1000',
         G == 'G1_DH',
         P %in% paste0('P=', c(2, 4, 8, 16)),
         parameter == 'E_Vg',
         L == 'L=1000',
         ancestral == 'Landrace',
         !(MD %in% c('DC', 'HC') & P == 'P=2')) %>% 
  ggplot(aes(x=MD, y = Mean))+
  geom_col(position = 'dodge', fill = cbPalette[2])+
  geom_text(y = 500, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_y_continuous(limits = c(0, 500))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  ylab(bquote('Estimated expectation for'~V[A]))+
  theme_light(base_size = 10)+
  theme(strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'none',
        axis.title.x = element_blank())

plot_tab_landrace <- ggplotGrob(landrace)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_landrace, 
                     "axis-b-1",
                     trim=FALSE)

landrace <- ggpubr::as_ggplot(tmp)

plot_grid(elite, landrace) 

ggsave('figures/fig2.jpg', width = 15, height = 7.5,
       bg = 'white', units = 'cm')


```

## Fig. 3

```{r}
gtable_filter_remove <- function (x, name, trim = TRUE){
    matches <- !(x$layout$name %in% name)
    x$layout <- x$layout[matches, , drop = FALSE]
    x$grobs <- x$grobs[matches]
    if (trim) 
        x <- gtable_trim(x)
    x
}

cross_df <- tibble(P_num = c(2, 4, 8, 16)) %>% 
  mutate(P = factor(P_num, levels = c(2, 4, 8, 16),
                    labels = paste0('P=', c(2, 4, 8, 16))),
         DC = P_num/2,
         FC = (P_num/2)**2,
         HC = (P_num * (P_num - 1)) / 2) %>% 
  pivot_longer(DC:HC, names_to = 'MD',
               values_to = 'crosses') %>% 
  mutate(MD = factor(MD,
                     levels = c('DC', 'FC', 'HC'))) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2'))

A <- summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         ancestral == 'Elite',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2')) %>% 
  ggplot(aes(x = MD))+
  geom_col(aes(y = Mean, fill = parameter), colour = 'black')+
  geom_text(y = 0.28, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 0.30))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0),
        legend.position = 'none')+
  ylab('Estimated variance of b')+
  labs(title = 'Elite')

plot_tab_A <- ggplotGrob(A)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_A, 
                     "axis-b-1",
                     trim=FALSE)

A <- ggpubr::as_ggplot(tmp)

B <- summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         ancestral == 'Landrace',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'),
         Mean = ifelse(P == 'P=2' & MD %in% c('DC', 'HC'), NA, Mean))  %>%
    filter(!(MD %in% c('DC', 'HC') & P == 'P=2')) %>% 
ggplot(aes(x = MD))+
  geom_col(aes(y = Mean, fill = parameter), colour = 'black')+
  geom_text(y = 0.28, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 0.30))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0),
        legend.position = 'none')+
  ylab('Estimated variance of b')+
  labs(title = 'Landrace')

plot_tab_B <- ggplotGrob(B)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_B, 
                     "axis-b-1",
                     trim=FALSE)

B <- ggpubr::as_ggplot(tmp)

legend <- get_legend(summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 16)),
         L == 'L=1000',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter,
                                   "var_Vg" = 'widehat(var(V[g]))',
                                   "var_Cb" = 'widehat(var(C[b]))', "var_Cw" = 'widehat(var(C[w]))',
                                   "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = N, fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 34000))+
  facet_grid(~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'bottom',
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0))+
  ylab(bquote(bar(sigma)[C]^2))+
  labs(title = 'EL '))

tmp <- plot_grid(plot_grid(A, B,
          nrow = 1,
          labels = LETTERS[1:2], align = 'v'), 
          legend, 
          ncol = 1,
          rel_heights = c(10, 1),
          label_size = 12)

tmp 

ggsave('figures/fig3.jpg', width = 15, height = 7.5, bg = 'white', units = 'cm')

summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         ancestral == 'Elite',
         parameter %in% c('var_bw')) %>% 
  select(P, MD, Mean) %>% 
  pivot_wider(values_from = Mean, names_from = MD) %>% 
  write.csv2(file = 'var_bw_fig3.csv')

```


## Fig. 4

Differences for N in G1 EL

```{r}
tmp <- summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=50', 'N=250', 'N=1000'),
         P %in% paste0('P=', c(2)),
         ancestral == 'Elite',
         L == 'L=1000',
         MD == 'FC',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = N, fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 0.2))+
  facet_grid(~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0))+
  ylab('Estimated variance of b')+
  labs(title = 'Elite')+
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))


tmp

ggsave('figures/fig4.jpg', width = 7.5, height = 7.5, bg = 'white', units = 'cm')


```

## Fig. 5

Difference between generations in EL

```{r}
tmp <- summary_df %>% 
    filter(N %in% c('N=50', 'N=250', 'N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         ancestral == 'Elite',
         MD == 'FC',
         L == 'L=1000',
         parameter %in% c('var_bw', 'var_bb')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = gsub('_', '-', G), fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  facet_grid(N~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(legend.title = element_blank(),
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank())+
  ylab('Estimated variance of b')+
  scale_y_continuous(limits = c(0, 0.21))+
  labs(title = 'Elite FC')

tmp 

ggsave('figures/fig5.jpg', width = 15, height = 20, bg = 'white', units = 'cm')


```


## Fig. 6

```{r}
cross_df <- tibble(P_num = c(2, 4, 8, 16)) %>% 
  mutate(P = factor(P_num, levels = c(2, 4, 8, 16),
                    labels = paste0('P=', c(2, 4, 8, 16))),
         DC = P_num/2,
         FC = (P_num/2)**2,
         HC = (P_num * (P_num - 1)) / 2) %>% 
  pivot_longer(DC:HC, names_to = 'MD',
               values_to = 'crosses') %>% 
  filter(MD %in% c('HC', 'DC'))

tmp <- summary_df %>% 
    filter(N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         ancestral == 'Elite',
         MD %in% c('HC', 'DC'),
         L == 'L=1000',
         parameter %in% c('var_bw', 'var_bb')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = gsub('_', '-', G)))+
  geom_col(aes(y = Mean, fill = parameter), colour = 'black')+
  geom_text(x = 'G1-DH', y = 0.29, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  facet_grid(MD~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(legend.title = element_blank(),
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank())+
  ylab('Estimated variance of b')+
  scale_y_continuous(limits = c(0, 0.31))+
  labs(title = 'Elite')

tmp

ggsave('figures/fig6.jpg', width = 15, height = 10, bg = 'white', units = 'cm')
```




# Supplement

## S1

Trait distribution

```{r}
summary_distribution <- simulation_df %>%
  mutate(Cw = Vg + Cw,
         Cb = Vg + Cb) %>% 
  select(P, `V[A]` = VA, `V[g]` = Vg,
         `V[g]+C[w]` = Cw, `V[g]+C[b]` = Cb) %>% 
  pivot_longer(-P, names_to = 'parameter') %>% 
  group_by(P, parameter) %>% 
    summarize(mean = round(mean(value)),
            var = round(var(value)),
            excess = round(kurtosis(value) - 3, 2),
            skewness = round(skewness(value), 2)) %>% 
  ungroup()

summary_distribution <- expectation_df %>% 
  select(P, E_Vg, var_Vg, var_Cw, var_Cb, var_VA) %>% 
  mutate(`Exp.mean_V[A]` = E_Vg,
         `Exp.var_V[A]` = var_VA,
         `Exp.mean_V[g]` = E_Vg,
         `Exp.var_V[g]` = var_Vg,
         `Exp.mean_V[g]+C[w]` = E_Vg,
         `Exp.var_V[g]+C[w]` = var_Vg + var_Cw,
         `Exp.mean_V[g]+C[b]` = E_Vg,
         `Exp.var_V[g]+C[b]` = var_Vg + var_Cb,
         .keep = 'unused') %>% 
  pivot_longer(-P) %>%
  mutate(value = round(value)) %>% 
  separate(name, 
           into = c('exp_var', 'parameter'),
           sep = '_') %>% 
  pivot_wider(names_from = exp_var, values_from = value) %>% 
  left_join(summary_distribution) %>% 
  mutate(parameter = factor(parameter, 
                            levels = c('V[g]', 'V[g]+C[w]',
                                       'V[g]+C[b]', 'V[A]')),
         P = recode_factor(P, 
                           'P=2' = "'P'=='2'",
                           'P=4' = "'P'=='4'"))


tmp <- simulation_df %>%
  mutate(Cw = Vg + Cw,
         Cb = Vg + Cb) %>% 
  select(P, `V[A]` = VA, `V[g]` = Vg,
         `V[g]+C[w]` = Cw, `V[g]+C[b]` = Cb) %>% 
  pivot_longer(-P, names_to = 'parameter') %>% 
  mutate(parameter = factor(parameter, 
                            levels = c('V[g]', 'V[g]+C[w]',
                                       'V[g]+C[b]', 'V[A]')),
         P = recode_factor(P, 
                           'P=2' = "'P'=='2'",
                           'P=4' = "'P'=='4'")) %>%
  ggplot(aes(x = value))+
  geom_histogram(bins = 100)+
  geom_label(data = summary_distribution,
             aes(label = paste0('mean = ', mean, ' (', Exp.mean, ')',
                                '\nvariance = ', var, ' (', Exp.var, ')',
                                '\nskewness = ', skewness,
                                '\nkurtosis = ', excess 
                                )),
             x = 900, y = 1800, size = 3)+
    facet_grid(parameter~P, labeller = 'label_parsed') +
  xlab('Realized values')+
  theme_light(base_size = 10)+
  theme(legend.position = 'none',
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        axis.ticks.x = element_blank())

tmp 

ggsave('figures/figS1.jpg', width = 15, height = 15, bg = 'white', units = 'cm')

```

## S2

```{r}
summary_distribution <- simulation_df %>%
  select(P, C, `C[w]` = Cw, `C[b]` = Cb,
         b, `b[w]` = bw, `b[b]` = bb) %>% 
  pivot_longer(-P, names_to = 'parameter') %>% 
  group_by(P, parameter) %>% 
    summarize(mean = round(mean(value), 2),
            var = round(var(value), 2),
            excess = round(kurtosis(value) - 3, 2),
            skewness = round(skewness(value), 2),
            prct_neg = round((sum(value < 0)/n()) * 100, 2)) %>% 
  ungroup() %>% 
  mutate(parameter = factor(parameter, 
                            levels = c('C', 'C[w]', 'C[b]',
                                       'b', 'b[w]', 'b[b]')),
         P = recode_factor(P, 
                           'P=2' = "'P'=='2'",
                           'P=4' = "'P'=='4'"))


A <- simulation_df %>%
  select(P, C, `C[w]` = Cw, `C[b]` = Cb,
         b, `b[w]` = bw, `b[b]` = bb) %>% 
  pivot_longer(-P, names_to = 'parameter') %>% 
  mutate(parameter = factor(parameter, 
                            levels = c('C', 'C[w]', 'C[b]',
                                       'b', 'b[w]', 'b[b]')),
         P = recode_factor(P, 
                           'P=2' = "'P'=='2'",
                           'P=4' = "'P'=='4'")) %>%
  filter(parameter %in% c('C', 'C[w]', 'C[b]')) %>% 
  ggplot(aes(x = value))+
  geom_histogram(bins = 100)+
  geom_label(data = filter(summary_distribution, parameter %in% c('C', 'C[w]', 'C[b]')),
             aes(label = paste0('mean = ', mean, 
                                '\nvariance = ', var, 
                                '\nskewness = ', skewness,
                                '\nkurtosis = ', excess,
                                '\nsign(-) = ', prct_neg, '%'
                                )),
             x = 600, y = 3000, size = 3)+
    facet_grid(parameter~P, labeller = 'label_parsed') +
  xlab('Realized values')+
  theme_light(base_size = 10)+
  theme(legend.position = 'none',
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        axis.ticks.x = element_blank())

B <- simulation_df %>%
  select(P, C, `C[w]` = Cw, `C[b]` = Cb,
         b, `b[w]` = bw, `b[b]` = bb) %>% 
  pivot_longer(-P, names_to = 'parameter') %>% 
  mutate(parameter = factor(parameter, 
                            levels = c('C', 'C[w]', 'C[b]',
                                       'b', 'b[w]', 'b[b]')),
         P = recode_factor(P, 
                           'P=2' = "'P'=='2'",
                           'P=4' = "'P'=='4'")) %>%
  filter(parameter %in% c('b', 'b[w]', 'b[b]')) %>% 
  ggplot(aes(x = value))+
  geom_histogram(bins = 100)+
  geom_label(data = filter(summary_distribution, parameter %in% c('b', 'b[w]', 'b[b]')),
             aes(label = paste0('mean = ', mean, 
                                '\nvariance = ', var, 
                                '\nskewness = ', skewness,
                                '\nkurtosis = ', excess,
                                '\nsign(-) = ', prct_neg, '%'
                                )),
             x = 2.5, y = 2500, size = 3)+
    facet_grid(parameter~P, labeller = 'label_parsed') +
  xlab('Realized values')+
  theme_light(base_size = 10)+
  theme(legend.position = 'none',
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank(),
        axis.ticks.x = element_blank())

plot_grid(A, B, labels = c('A', 'B'), ncol = 1)

ggsave('figures/figS2.jpg', width = 15, height = 20, bg = 'white', units = 'cm')
```



## S3

Histograms of MAF

```{r}
full_pop <-  readRDS('input/founding_pop.RData')

EL = do.call(cbind, lapply(full_pop$haplotypes, function(x){
  x[grepl('Elite',rownames(x)),]/2
}))

LR = do.call(cbind, lapply(full_pop$haplotypes, function(x){
  x[grepl('Landrace',rownames(x)),]/2
}))

tmp <- tibble(Landrace = colMeans(LR),
       Elite = colMeans(EL)) %>% 
  pivot_longer(everything()) %>% 
  mutate(MAF = ifelse(value >= 0.5, 1-value, value)) %>% 
  ggplot(aes(x = MAF, fill = name))+
  geom_density(alpha = 0.5, adjust = 0.5)+
  scale_fill_manual(values = cbPalette)+
  theme_light(base_size = 10)+
  theme(panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'bottom')+
  ylab('Density')+
  xlab('Minor allele frequency')

tmp

ggsave('figures/figS3.jpg', width = 7.5, height = 7.5, bg = 'white', units = 'cm')
```


## S4

```{r}
gtable_filter_remove <- function (x, name, trim = TRUE){
    matches <- !(x$layout$name %in% name)
    x$layout <- x$layout[matches, , drop = FALSE]
    x$grobs <- x$grobs[matches]
    if (trim) 
        x <- gtable_trim(x)
    x
}

cross_df <- tibble(P_num = c(2, 4, 8, 16)) %>% 
  mutate(P = factor(P_num, levels = c(2, 4, 8, 16),
                    labels = paste0('P=', c(2, 4, 8, 16))),
         DC = P_num/2,
         FC = (P_num/2)**2,
         HC = (P_num * (P_num - 1)) / 2) %>% 
  pivot_longer(DC:HC, names_to = 'MD',
               values_to = 'crosses') %>% 
  mutate(MD = factor(MD,
                     levels = c('DC', 'FC', 'HC'))) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2'))

A <- summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         ancestral == 'Elite',
         parameter %in% c('var_Vg', 'var_Cb', 'var_Cw')) %>%
  mutate(parameter = recode_factor(parameter, 
                                   "var_Cb" = 'widehat(var(C[b]))',
                                   "var_Cw" = 'widehat(var(C[w]))',
                                   "var_Vg" = 'widehat(var(V[g]))')) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2')) %>%
  ggplot(aes(x = MD))+
  geom_col(aes(y = Mean, fill = parameter), colour = 'black')+
  geom_text(y = 42000, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_fill_manual(values = c(cbPalette[c(5, 4)], 'magenta'), labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 45000))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0),
        legend.position = 'none')+
  ylab(bquote('Estimated variance of'~V[A]))+
  labs(title = 'Elite')

plot_tab_A <- ggplotGrob(A)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_A, 
                     "axis-b-1",
                     trim=FALSE)

A <- ggpubr::as_ggplot(tmp)

B <- summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         ancestral == 'Landrace',
         parameter %in% c('var_Vg', 'var_Cb', 'var_Cw')) %>%
  mutate(parameter = recode_factor(parameter, 
                                   "var_Cb" = 'widehat(var(C[b]))',
                                   "var_Cw" = 'widehat(var(C[w]))',
                                   "var_Vg" = 'widehat(var(V[g]))')) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2')) %>%
  ggplot(aes(x = MD))+
  geom_col(aes(y = Mean, fill = parameter), colour = 'black')+
  geom_text(y = 42000, aes(label = crosses), data = cross_df, size = 2.5)+
  scale_fill_manual(values = c(cbPalette[c(5, 4)], 'magenta'), labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 45000))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0),
        legend.position = 'none')+
  ylab(bquote('Estimated variance of'~V[A]))+
  labs(title = 'Landrace')

plot_tab_B <- ggplotGrob(B)

tmp <- plot_filtered <- gtable_filter_remove(plot_tab_B, 
                     "axis-b-1",
                     trim=FALSE)

B <- ggpubr::as_ggplot(tmp)


legend <- get_legend(summary_df %>%
  filter(G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 16)),
         L == 'L=1000',
         parameter %in% c('var_Vg', 'var_Cb', 'var_Cw')) %>%
  mutate(parameter = recode_factor(parameter,
                                   "var_Cb" = 'widehat(var(C[b]))', "var_Cw" = 'widehat(var(C[w]))',
                                   "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))',
                                   "var_Vg" = 'widehat(var(V[g]))')) %>% 
  filter(!(MD %in% c('DC', 'HC') & P == 'P=2')) %>%
  ggplot(aes(x = N, fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = c(cbPalette[c(5, 4)], 'magenta'), labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 34000))+
  facet_grid(~P, switch = 'x', scales = 'free_x', space = 'free_x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = 'bottom',
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0))+
  ylab(bquote(bar(sigma)[C]^2))+
  labs(title = 'EL '))

tmp <- plot_grid(plot_grid(A, B,
          nrow = 1,
          labels = LETTERS[1:2], align = 'v'), 
          legend, 
          ncol = 1,
          rel_heights = c(10, 1))

tmp 

ggsave('figures/figS4.jpg', width = 15, height = 7.5, bg = 'white', units = 'cm')


```

## S5

Differences between number of QTL

```{r}
gtable_filter_remove <- function (x, name, trim = TRUE){
    matches <- !(x$layout$name %in% name)
    x$layout <- x$layout[matches, , drop = FALSE]
    x$grobs <- x$grobs[matches]
    if (trim) 
        x <- gtable_trim(x)
    x
}

A <- summary_df %>% 
  filter(P != 'P=2',
         G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         ancestral == 'Elite',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = L, fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 0.28))+
  facet_grid(MD~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0),
        axis.title.y = element_blank())+
  ylab('Estimated variance of b')+
  labs(title = ' ')

A

B <- summary_df %>% 
  filter(P == 'P=2',
         G == 'G1_DH',
         N %in% c('N=1000'),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         ancestral == 'Elite',
         parameter %in% c('var_bb', 'var_bw')) %>%
  mutate(parameter = recode_factor(parameter, "var_bb" = 'widehat(var(b[b]))', "var_bw" = 'widehat(var(b[w]))'))  %>%
  ggplot(aes(x = L, fill = parameter))+
  geom_col(aes(y = Mean), colour = 'black')+
  scale_alpha_manual(values = c(0.5, 1))+
  scale_fill_manual(values = cbPalette[c(7, 6)], labels = scales::parse_format())+
  scale_y_continuous(limits = c(0, 0.28))+
  facet_grid(MD~P, switch = 'x')+
  theme_light(base_size = 10)+
  theme(axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.y = element_blank(),
        legend.position = 'none')+
  ylab('Estimated variance of b')+
  labs(title = 'Elite N=1000')


plot_tab <- ggplotGrob(B)

plot_filtered <- gtable_filter_remove(plot_tab, 
                     c("panel-2-1","panel-1-1",
                       "axis-l-1", "axis-l-2"),
                     trim=FALSE)

B <- ggpubr::as_ggplot(plot_filtered)

plot_grid(B, A, ncol = 2,
          rel_widths = c(1, 3),
          axis = 'b')

ggsave('figures/figS5.jpg', width = 15, height = 12, bg = 'white', units = 'cm')

```

## S6

non-linear regression

```{r}
reg_df <- summary_df %>%
  filter(parameter == 'var_bb') %>%
  mutate(N_num = as.numeric(as.character(gsub('N=', '', N)))) %>%
  select(MD, ancestral, L, P, N, N_num, G, Mean) %>%
  mutate(bla = 'real') %>%
  unite(col = 'G', c(bla, G), sep = '_') %>%
  pivot_wider(names_from = G, values_from = Mean) %>%
  mutate(start = `real_G1_DH`) %>%
  pivot_longer(-c(MD, ancestral, L, P, N, N_num, start)) %>%
  separate(name, into = c('type', 'G', NA), sep = '_') %>%
  group_by(MD, ancestral, L, P, N, N_num) %>%
  mutate(
    rec_steps = 0:3,
    omega_factor = vapply(0:3, function(r) {
      sum(0.25 ** (0:r))
    }, FUN.VALUE = numeric(1))
  ) %>%
  ungroup() 

model = function(df){
  nls(value ~ (start - omega) * 0.25**(rec_steps) + ((4 - 1/4**rec_steps)/3) * omega, start = c(omega=0), data = df)}

summary_reg <- reg_df %>%
  filter(rec_steps > 0) %>% 
  group_by(MD, ancestral, L, P, N) %>% 
  nest() %>% 
  mutate(model = map(data, model),
         tmp = map(model, augment),
         summary = map(model, tidy)) %>% 
  unnest(summary)

reg_plot <- reg_df %>%
  filter(rec_steps > 0) %>% 
  group_by(MD, ancestral, L, P, N) %>% 
  nest() %>% 
  mutate(model = map(data, model),
         tmp = map(model, augment),
         summary = map(model, tidy)) %>% 
  unnest(tmp) %>% 
  group_by(MD, ancestral, L, P) %>% 
  mutate(rec_steps = rep(1:3, 3)) %>% 
  ungroup() %>% 
  left_join(reg_df %>%
  # filter(rec_steps > 0) %>% 
  mutate(expectation = start * (1/4)**(rec_steps))) %>% 
  pivot_longer(c(value, .fitted, expectation)) %>% 
  bind_rows(reg_df %>% 
  mutate(expectation = value *(1/4)**rec_steps) %>% 
  pivot_longer(c(value, expectation)) %>% 
  filter(rec_steps == 0,
         name == 'value')) %>% 
  mutate(name = factor(name, 
                       levels = c('value', '.fitted', 'expectation'))) 

tmp <- reg_plot %>% 
  filter(ancestral == 'Elite', 
         N_num %in% c(50, 250, 1000),
         P %in% paste0('P=', c(2, 4, 8, 16)),
         L == 'L=1000',
         MD == 'FC') %>% 
  ggplot(aes(paste0(G, '-DH'), value, shape = name))+
  geom_point(alpha = 0.5, size = 2)+
  facet_grid(N~P, switch = 'x')+
  scale_shape_manual(values = c(1, 2, 8))+
  scale_y_continuous(limits = c(0, 0.17))+
  theme_light(base_size = 10)+
  theme(strip.text.y = element_text(angle = 0),
        legend.position = 'none',
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ylab(bquote('Estimated variance of'~b[b]))+
  xlab('Number of additional recombination steps')+
  labs(title = 'Elite FC')

tmp

ggsave('figures/figS6.jpg', width = 15, height = 12, bg = 'white', units = 'cm')
```

## S7

```{r}
A <- p_df %>% 
  pivot_longer(-y) %>% 
  filter(value > 0,
         name == 'p=0.5') %>% 
  ggplot(aes(x = y, y = value, colour = y < 0))+
  ylab(bquote("P[Y = y|L = 1000," ~ beta ~ "= 0.5]"))+
  geom_segment(aes(xend = y), yend = 0)+
  scale_colour_manual(values = c('cornflowerblue', 'orange'))+
  geom_hline(yintercept = 0)+
  theme_light(base_size = 10)+
  theme(legend.position = 'none',
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank())+
  coord_cartesian(xlim = c(-600, 7500))

A 

PY_y <- p_df %>% 
  pivot_longer(-y) %>% 
  filter(value > 0,
         name == 'p=0.5') %>% 
  pull(value)

B <- p_df %>% 
  pivot_longer(-y) %>% 
  filter(value > 0,
         name == 'p=0.5') %>% 
  mutate(cumulative = cumsum(value)) %>% 
  ggplot(aes(x = y, y = cumulative, colour = y < 0))+
  ylab(bquote("P[Y" <= "y|L = 1000," ~ beta ~ "= 0.5]"))+
  geom_segment(aes(xend = y), yend = 0)+
  scale_colour_manual(values = c('cornflowerblue', 'orange'))+
  geom_hline(yintercept = 0)+
  theme_light(base_size = 10)+
  theme(legend.position = 'none',
        strip.text.y = element_text(angle = 0),
        panel.grid = element_blank())+
  coord_cartesian(xlim = c(-600, 7500))

B

plot_grid(A, B, ncol = 2, labels = c('A', 'B'))

ggsave('figures/figS7.jpg', width = 15, height = 7.5, bg = 'white',
       units = 'cm')

```

## S8

```{r}

```


## S9

Histogram of the D values in the ancestral populations

```{r}
D_ancestral %>%
  mutate(elements = factor(elements, 
                           levels = c('diag',
                                      'within_chr',
                                      'between_chr',
                                      'off_diag'),
                           labels = c('V',
                                      'W',
                                      'B',
                                      'W & B'))) %>% 
  filter(elements != 'V') %>% 
  ggplot(aes(value))+
  geom_histogram(bins = 100)+
  facet_grid(elements~population, scales = 'free_y')+
  theme_light(base_size = 10)+
  theme(strip.text.y = element_text(angle = 0),
        panel.grid = element_blank())+
  xlab(bquote(d[ij]))

  ggsave('figures/figS9.jpg', width = 15, height = 15, bg = 'white',
       units = 'cm')
```




## Tab. S1

AMOVA

```{r}
library(ade4)
full_pop <-  readRDS('input/founding_pop.RData')

EL = do.call(cbind, lapply(full_pop$haplotypes, function(x){
  x[grepl('Elite',rownames(x)),]
}))

LR = do.call(cbind, lapply(full_pop$haplotypes, function(x){
  x[grepl('Landrace',rownames(x)),]
}))

within_EL = (sum(as.matrix(dist(EL/2, diag = TRUE, upper = TRUE))**2)/ (2*nrow(EL))) / (nrow(EL) - 1)

within_LR = (sum(as.matrix(dist(LR/2, diag = TRUE, upper = TRUE))**2)/ (2*nrow(LR))) / (nrow(LR) - 1)

groups = tibble(EL = c(rep(1, nrow(EL)), rep(0, nrow(EL))),
                LR = c(rep(0, nrow(LR)), rep(1, nrow(LR))))
dist_mtx = dist(rbind(EL/2, LR/2))

mean(c(within_LR, within_EL))

amova(groups, dist_mtx)

within_EL
within_LR

randtest(amova(groups, dist_mtx),
         nrepet = 10000, alter = 'two-sided')
```

## Tab. S2

```{r}
reg_df <- summary_df %>%
  filter(parameter == 'var_bb') %>%
  mutate(N_num = as.numeric(as.character(gsub('N=', '', N)))) %>%
  select(MD, ancestral, L, P, N, N_num, G, Mean) %>%
  mutate(bla = 'real') %>%
  unite(col = 'G', c(bla, G), sep = '_') %>%
  pivot_wider(names_from = G, values_from = Mean) %>%
  mutate(start = `real_G1_DH`) %>%
  pivot_longer(-c(MD, ancestral, L, P, N, N_num, start)) %>%
  separate(name, into = c('type', 'G', NA), sep = '_') %>%
  group_by(MD, ancestral, L, P, N, N_num) %>%
  mutate(
    rec_steps = 0:3,
    omega_factor = vapply(0:3, function(r) {
      sum(0.25 ** (0:r))
    }, FUN.VALUE = numeric(1))
  ) %>%
  ungroup() 

model = function(df){
  nls(value ~ (start - omega) * 0.25**(rec_steps) + ((4 - 1/4**rec_steps)/3) * omega, start = c(omega=0), data = df)}

summary_reg <- reg_df %>%
  filter(rec_steps > 0) %>% 
  group_by(MD, ancestral, L, P, N) %>% 
  nest() %>% 
  mutate(model = map(data, model),
         tmp = map(model, augment),
         summary = map(model, tidy)) %>% 
  unnest(summary)

summary_reg %>% 
  filter(L == 'L=1000',
         ancestral == 'Elite',
         MD == 'FC') %>% 
  ungroup() %>% 
  select(P, N, estimate, std.error) %>% 
  pivot_wider(names_from = N, values_from = estimate:std.error) %>% 
  mutate_at(2:7, round, digits = 3)
```

## GPD decay

```{r}
summary_reg_df <- bind_rows(list(Landrace = landrace,
                         Elite = elite), .id = 'ancestral') 

fitcurve_plot <- function(x,p,n = 115) {
  ((10 + p*x)) / ((2+p*x) * (11 + p*x) ) *
    ( 1 + ( (3+ p*x) * (12 + 12 * p + p^2*x^2)) / ( n*(2+p*x) * (11 + p*x)))	
}


plot_df <- bind_rows(list(Landrace = bind_rows(GPD_LR, .id = 'chr'),
               Elite = bind_rows(GPD_EL, .id = 'chr')), 
          .id = 'ancestral') %>% 
  filter(!is.na(r2),
         distance <= 1) 
  

p <- ggplot(plot_df, aes(distance, r2))+
  geom_point(alpha = 0.01)
  
for(i in 1:nrow(summary_reg_df)){
p <- p + geom_function(data = transform(plot_df,
                               ancestral = summary_reg_df$ancestral[i],
                               chr = summary_reg_df$chr[i]),
              fun = fitcurve_plot,
              args = list(p = summary_reg_df$p[i]),
              colour = 'red')
}

p+
  geom_hline(yintercept = 0.1, colour = 'blue')+
  geom_label(data = summary_reg_df, 
             aes(label = paste0(round(r2_decay * 100, 2),
                                ' cM')),
             x = 0.75, y = 0.75, size = 2.5)+
  facet_grid(ancestral~factor(chr, levels = 1:10, labels = paste0('chr:', 1:10)),
             scales = 'free_x', space = 'free_x')+
  scale_x_continuous(labels = function(x){x*100})+
  xlab('Distance in cM')+
  ylab(bquote(r^2))
```



# Preparatory analysis

The analysis shows that while with 250 reps and 100 sets the maximum SE/Mean is minimal (3.2%), the median is nearly twice as high as compared to 500 reps and 50 sets (13.6% and 7.8%). Our choice of numbers replications and sets of QTL positions was based on this.

```{r}
prep_df <- readRDS('output/preparatory_summary_df.RData')

summary(prep_df$rep)
summary(prep_df$set)

prep_df %>% 
  filter(set == 0) %$%
  table(parameter)


purrr::map2(c(50, 100, 250, 500), c(500, 250, 100, 50), function(x, y){
prep_df %>% 
  mutate(SE_new = sqrt(rep/x + set/y + Residual/(x * y))) %>% 
  filter(L == 'L=1000') %$% 
  summary(SE_new/Mean) 
})

purrr::map2(c(50, 100, 250, 500), c(500, 250, 100, 50), function(x, y){
prep_df %>% 
  mutate(SE_new = sqrt(rep/x + set/y + Residual/(x * y))) %$% 
  summary(SE_new/Mean) 
})
```

